Scaling rules for the ionization of biological molecules by highly charged ions

### 2020-06-02 scalings/review/response.tex

\documentclass[a4paper,12pt]{article}
\usepackage[utf8]{inputenc}
\setlength{\parindent}{0pt}
\usepackage{soul}
\usepackage{xcolor}
\usepackage[margin=0.8in]{geometry}
\usepackage{graphicx}

\setstcolor{blue}
\def\reviewer#1{\vspace{0.35cm}\textsl{#1}}
\def\reply#1{\vspace{0.1cm}\textcolor{blue}{#1}}

\title{Response to referee reports \\ JPHYSB-106014}

\begin{document}
\maketitle

\reply{We appreciate the comments and questions of the three referees. 
They were taken into account, as explained in the replies to each 
referee. In what follows, the referee reports are in black letters and 
our answers and comments in blue. }

\reply{\textbf{Summary of the modifications:} The manuscript was 
changed from letter to regular manuscript by the editorial board. 
Attending to this fact and the comments by the three referees 
requesting further explanations, we decided to modify the manuscript by
dividing the text into three sections: Introduction, Scaling rules (with
subsections), and Conclusions. Additional specifications of the scaling 
models are given in Section II-A, B, and C. Figure 2 was improved by 
including a gray area of $\pm 30\%$ that covers most of the available 
data. Also, experiments for H$^+$ and He$^{+2}$ in water (missing in 
the previous version) were added to the figures (Refs. 25, 26, and 27), 
extending the energy range where the present scaling agrees with the 
experimental data. 
% the agreement of the present scaling for H$^+$ in water above 40 keV 
% and He$^{+2}$ in water up to 10 MeV/amu. 
Following the suggestion by one of the referees, we changed the title 
from ``Universal scaling for the ionization...'' to ``Scaling 
rules for the ionization...''}.

% ----------------------------------------------------------------------
\section{Reply to report of the First Referee}
% ----------------------------------------------------------------------

\reviewer{The authors propose the universal scaling law of the 
ionization cross sections by the impact of multiply charged atomic ions 
on the basis of the experimental data reported in their very recent 
publication [J. Phys. B, 53, 055201 (2020)]. Their conclusion of this 
newly submitted paper is that they found the scaling law of Eq. (1) 
with the alpha value of 1.2. However, as the authors mention in the 
introduction part, the same scaling law with the alpha value of 
4/3 = 1.333 was reported in 2013 by DuBois et al. [Ref. 15]. The 
authors should explain if the scaling with alpha = 4/3, which is very 
close to their value of 1.2, could not explain the experimental data 
they recorded. If the value of 1.2 is much better than 1.333, the 
author should explain possible reason why alpha should be 1.2 on the 
basis of the physical meaning of the parameter alpha.}

\vspace{0.2cm}

\reply{The value of $\alpha=1.2$ was found numerically as the best
convergence of the CDW-SSM cross sections of the forty collisional 
systems over the broadest energy range. Our methodology does not rely 
on the fitting of experimental data. The work considers the scaling of 
our theoretical results, which is tested afterward with the 
experimental data. We also noticed that all the maxima of the 
theoretical cross sections scaled with $\alpha=1.2$ lay at the same 
value of $E/Z^{2-\alpha}=E/Z^{0.8}$. In the present version of our 
manuscript, we modified lines 59-61 to clarify our choice of $\alpha$.}

\reply{About the  physical meaning of this parameter, the scaling by
Janev et~al.~[14] uses $E/Z^1$ as abscissa based on the long-distance
approximation for the ionization cross sections. Similarly, DuBois 
et~al.~[15] propose $E/Z^{2/3}$. The 2/3 in Ref.~[15] was found by 
trial and error. However, the dependence with $Z^{2/3}$ recalls the 
Thomas Fermi model for a constant density of electrons around the ion. 
% Also in [15], it is mentioned that $E/Z^{0.88}$ joins  the maximum of 
% the cross sections of He and H$_2$. 
In either case, both models are approximations, and our numerical 
proposal of $\alpha$ is located between them.}

\vspace{0.2 cm}
% ----------------------------------------------------------------------
\section{Reply to report of the Second Referee}

The authors propose a scaling formula for cross sections for ionization
of complex molecules by multiply charged ions. Such cross sections are
required, e.g., for estimating radiation damage in biomolecules. A
routine ab-initio calculation of such cross sections by quantum methods
is beyond reach of the state-of-the-arts methods. In this situation, 
it is helpful to have semi-empirical scaling formulas for modeling
purposes. Thus, the paper is timely and serves a need. It is concise.
However, there are some language issues. The paper should eventually be
published as a JPB letter after the authors having adequately responded
to my detailed points below.

% ----------------------------------------------------------------------

\vspace{0.2cm}

\reply{The points 1)-4), 6) and 7) raised by the referee are language
corrections. They were all considered and modified in the manuscript
where appropriate.}

%\reviewer{1) page1, line 11: `` ... from +1 to +8 colliding with five ...''}
%\reviewer{2) page 1, line 14: `` ... fourty ...''}
%\reviewer{3) page 1, line 26: `` ... continuum distorted-wave ...''}
%\reviewer{4) page 1, line 42: ``... first order approximation in experimental ...''}
%\reviewer{6) page 1, line 20, column 2: ``... fourty ...''}
%\reviewer{7) page 1, line 24,, column 2: `` ... and five ion species: ...''}

\reviewer{5) page 1, line 53: I do not see how any value for alpha keeps
the $Z^2/E$ relationship? $\alpha = 1.2$ results in $Z^{0.8}/E$.}

\reply{In first order approaches, the ionization cross section $\sigma$
depends linearly with $Z^2/E$. This dependence is valid in the first Born 
approximation, and at high impact energies, $\sigma/Z^{2}$ scales with
$E$ for any charged ion. Within this approximation, $\sigma/Z^{\alpha}$
will be a function of $Z^{2-\alpha}/E$ for any value of $\alpha$. At
intermediate energy, the linear dependence with $Z^2/E$ no longer holds,
so we numerically sought for the value of $\alpha$ that best merges
the cross sections for different impinging charges.}

\reviewer{8) Figure captions. Both caption should mention that all 
plots are for alpha=1.2.}

\reply{Done.}

\reviewer{9) page 2, line 40, column 2: How broad is the ``narrow 
band''? The authors should supply some estimate for the uncertainty of 
the scaling formula. I consider this very important.}

\reply{We estimated the uncertainty of the scaling formula to 20\% for 
the theoretical curves of the different targets, and 30\% to incorporate 
most of the available data. The gray area drawn in Fig.~2 accounts for 
this uncertainty. We also included a sentence to refer to this subject 
in lines 119-122.}

\vspace{0.2 cm}
% ----------------------------------------------------------------------
\section{Reply to report of the Third Referee}
% ----------------------------------------------------------------------

This paper tries to obtain a scaling law for ionization of some 
biological molecules by highly charged ions. As is well known, such
processes are quite relevant and important in medical physics. Since 
the molecules considered are complex, formal theories of ionization are
difficult to apply and scaling laws, if found applicable, greatly 
simplify the process of obtaining reliable cross sections. In this
respect the present paper is quite significant. The authors have
proposed a scaling law and have applied it to several bio molecules 
and have tried to extend and justify the validity of the scaling law 
by including results for N2, O2, CO, CO2 and CH4 to show they have the
same scaling behaviour. However, there are several aspects of the paper
which I either do not understand clearly, or have reservations.

\reviewer{1. The authors mention $\sigma$ to be a function of $E/Z$. 
But the functional form is nowhere mentioned in the paper. What is the
functional dependence on $E/Z$ the authors consider in the work?}

\reply{In different energy regions, the ionization cross section 
follows different $Z/E$ laws. At high energies, it is a function of 
$Z^2/E$, as predicted by the first Born approximation. The CDW 
approximation has a more complex dependence with $Z/E$ but tends to the
Born approximation at high energies, as shown in our previous work [8]. 
The objective of this manuscript is precisely to deepen into the study 
of the dependence of the ionization cross sections with $Z$ and $E$ in 
the intermediate energy region. Janev and Presnyakov [14], in their 
paper from 1980, suggest that $\sigma/Z$ is a linear function of $E/Z$ 
at intermediate energies (around the maximum of the cross sections).}

\reply{In this version, we further explained the models and cross 
section dependence with $E/Z$ (line 24 and the second paragraph of 
Section II A).}

\reviewer{2. I have reservations in calling the scaling law developed 
a universal scaling law. Not only have the authors considered only 8
molecular targets and five incident ions, the results, as the authors
claim, are applicable only around the maximum of the cross sections. 
Though this may be reasonable for practical purposes, in my opinion, 
the results are in too narrow range for calling it a universal scaling
law. I would recommend dropping the word universal everywhere in the
paper.}

\reply{We called our scaling ``universal'' in the sense that it can be
implemented for any ion-target combination within the intermediate 
energy range, i.e., $E/Z^{0.8} \simeq (0.04-5)$ MeV/amu. Nevertheless, 
we followed the referee's recommendation and dropped the word 
``universal'' from the title and the rest of the manuscript. Instead, 
we emphasized the independent nature of the reduced scaling law for 
different ions and molecular targets. We also remaked that even 
the experimental data (in the energy range where our theoretical method
starts to fail) follows the scaling proposed. In that sense, we changed
the last sentence of the abstract, Section II C (lines 127-130, 136),
and the Conclusions (lines 140-154, 157-160).}

\reviewer{3. Figure 1 somehow needs to be improved or made larger,
because it is very difficult to gauge the deviation of the experimental
points from the curves. Is it possible to separate the figure in to two
figures (eg. top four Figure 1, bottom four Figure 2)?}

\reply{Separating Fig.~1 would complicate the ratio of the figure; 
instead, we considered convenient to make the figure larger to utilize 
the space better. We have also reduced the scale of the y-axis of the 
subplots to improve the visualization of the data.}

\reviewer{4. Figure captions need to be significantly improved as it is
difficult to read and correlate the legend descriptions with those in
the figure. For the caption in Figure 2, it is not clear what the
CDW-SSM theoretical results represent ? What is the target molecule
considered?}

\reply{We believe the confusion about Fig.~2 is originated from 
inconsistencies in the text. Figure~2 shows a scaling that is 
independent of the ion charge of the projectile and the molecular 
target. All the figures have the same CDW-SSM results as inputs. We 
improved the explanation about our calculations by adding the last 
paragraph of Section II B (and Eq. (2)), and the first sentence of 
Section II. C. The notation of Eq. (3) was slightly changed to underline 
that the starting point is the CDW-SSM cross sections for molecules.}

%\reviewer{5. The language should be considerably improved as there are many unusual statements.}

%\reviewer{6. There are many typographical errors.}

\vspace{0.5cm}
\reply{Regarding points 5) and 6), we revised our manuscript and we 
believe it has been considerably improved.}

\end{document}

### 2020-06-02 scalings/scalings.tex
cleaned folder

\definecolor{olive}{RGB}{182,187,37}
\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

@@

\title[Scaling rules for the ionization of biological
molecules]{Scaling rules for the ionization of biological
molecules by highly charged ions}
\author{A. M. P. Mendez, C. C. Montanari, J. E. Miraglia}
\affiliation{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio 
(CONICET-UBA), \\ Buenos Aires, Argentina.}

\date{\today}% It is always \today, today,

\begin{abstract}
We investigate scaling rules for the ionization cross sections of
multicharged ions on molecules of biological interest. The cross 
sections are obtained using a methodology presented in [Mendez 
\textit{et al.} J. Phys B (2020)], which considers distorted-wave 
calculations for atomic targets combined with a molecular 
stoichiometric model. We examine ions with nuclear charges $Z$ from 
$+1$ to $+8$ impacting on five nucleobases --adenine, cytosine, 
guanine, thymine, uracil--, tetrahydrofuran, pyrimidine, and water. 
We investigate %the scaling 
scaling rules of the ionization cross section with 
the ion charge and the number of active electrons per molecule. 
%valid in the intermediate to 
%high energy range, i.e., 0.2-5 MeV/amu for oxygen impact. 
% Furthermore, 
% we introduce a modified scaling rule, which takes into account the 
% number of active electrons per molecule. 
Combining these two features, we define a scaling law  for any ion and molecular target %independent of  the ion charge and the complexity of the molecular target
, which is 
valid in the intermediate to high energy range, i.e., 0.2-5 MeV/amu for oxygen impact. %Then, 
Thus, the forty 
ion-molecule systems analyzed here can be merged into a single band. 
We confirm the generality of our independent scaling law with several
collisional systems. 
\end{abstract}

@@

\maketitle
%\ioptwocol
\linenumbers

@@

This contribution constitutes a follow-up of our previous work~\cite{MendezJPB20}. We 
introduce here a two-folded scaling rule for the ionization cross 
sections of complex molecules by charged ions. Our approach considers 
the dependence of the cross section with the ion charge $Z$ 
and incorporates the scaling of the ionization with the number of 
active electron $n_e$ of the molecular targets. Scaling rules are 
generally very useful since they can be used as first-order 
approximations in experimental measurements and multipurpose codes.

@@

\begin{figure*}[!htb]
\centering
\includegraphics[width=0.87\textwidth]{zscale_alpha.eps}
\caption{(Color online) Scaled ionization cross section $\sigma/Z^{\alpha}$ 
as a function of ion impact energy $E/Z^{2-\alpha}$ with $\alpha=1.2$. 
Colors are associated with the incident ion labeled on top of the figure. 
Curves: present CDW-SSM theoretical results. Symbols: experimental 
impact of \mbox{\LARGE$\color{red}\circ$} H$^+$~\cite{iriki2011} and
{\fontsize{11}{20}$\color{Purple}\pmb{\lozenge}$} C$^{+6}$ 
\cite{tribedi2019} on adenine;
H$^+$ on {\fontsize{11}{20}$\color{red}\pmb{\triangle}$} uracil~\cite{itoh2013}, 
{\fontsize{11}{20}$\color{NavyBlue}\pmb{\ominus}$} C$^{+4}$, 
{\fontsize{11}{20}$\color{Purple}\pmb{\oplus}$} C$^{+6}$, O$^{+6}$, F$^{+6}$, and 
{\fontsize{11}{20}$\color{BurntOrange}\pmb{\otimes}$} O$^{+8}$, 
F$^{+8}$ on uracil~\cite{agnihotri2012,agnihotri2013};
H$^+$ on {\fontsize{11}{20}$\color{red}\pmb{\triangledown}$} 
pyrimidine~\cite{wolff2014}, and 
{\fontsize{10}{20}$\color{red}\pmb{\meddiamond}$} THF~\cite{wang2016};
% water
\mbox{\fontsize{11}{20}$\color{red}\pmb{\odot}$}~\cite{Luna2007}, 
{\fontsize{11}{20}\color{red}$\pmb{\logof}$}~\cite{Rudd86}, 
{\fontsize{11}{20}\color{red}$\pmb{\varowedge}$}~\cite{pRudd85}, 
{\fontsize{11}{20}\color{red}$\pmb{\varoslash}$}~\cite{toburen80} H$^+$,
{\fontsize{11}{20}\color{ForestGreen}$\pmb{\varolessthan}$}~\cite{Ohsawa05},
{\fontsize{11}{20}\color{ForestGreen}$\pmb{\varogreaterthan}$}~\cite{Rudd85},
{\fontsize{11}{20}\color{ForestGreen}$\pmb{\varoslash}$}~\cite{toburen80} He$^{+2}$,
{\fontsize{11}{20}\color{Purple}$\pmb{\ovee}$}~C$^{+6}$ \cite{DalCappello2009,Bhattacharjee17}, and 
{\fontsize{11}{20}\color{BurntOrange}$\pmb{\obslash}$}
O$^{+8}$ \cite{Tribedi_O_water} on water.
Markers~$\square$~\cite{rahman2016}, 
$\rhd$~\cite{bug2017}, 
$\lhd$~\cite{wolf2019}, and 
$\medstar$~\cite{fuss2009} correspond to electron impact ionization with 
the equi-velocity conversion.}
\label{fig:zreduced}
\end{figure*}

We found two types of $Z$-scaling laws in the literature applicable 
to the intermediate impact energy range. The rule suggested by Janev 
and Presnyakov~\cite{janev1980} 
%depends linearly with ion charge $Z$, considering 
considers $\sigma/Z$ versus $E/Z$ to be the \textit{natural} reduced  
form of the ionization cross section $\sigma$ and the incident ion energy $E$. More recently, Montenegro and co-workers~\cite{dubois13,
montenegro_pra13} suggested an alternative scaling by taking into 
account that the cross section is a function of $Z^2/E$ at high 
energies. Their scaling, given by 
\begin{equation}
 \sigma/Z^{\alpha}=f(E/Z^{2-\alpha}),
\label{eq:Montenegro}
\end{equation}
keeps the $Z^2/E$ relationship for any value of the parameter 
$\alpha$. The authors proposed $\alpha=4/3$ for the ionization of He and 
H$_2$ by differently charged ions~\cite{dubois13}.

@@

We also examined the experimental data available for the forty 
ion-target systems~\cite{itoh2013,iriki2011,wolff2014,wang2016,
tribedi2019,agnihotri2012,agnihotri2013,Luna2007,Rudd86,Rudd85,pRudd85,
toburen80,Ohsawa05,Bhattacharjee17,Luna_Li_water,DalCappello2009,
Tribedi_O_water} with the $Z^\alpha$-scaling rule. For targets with none or little experimental data, we included
electron impact ionization results~\cite{rahman2016,bug2017,wolf2019,
fuss2009} at high velocity with the corresponding equivelocity 
conversion.
As can be noted, 
most of the data in Fig.~\ref{fig:zreduced} confirm the present 
scaling, even for O$^{+8}$ in water~\cite{Tribedi_O_water}. 
Only two data sets are off our predictions: the ionization cross sections of 
uracil by swift C, O, and F ions from 
Refs~\cite{agnihotri2012,agnihotri2013}% are too low  compared with our CDW-SSM results
; and the values 
for Li$^{+3}$ in water from Ref.~\cite{Luna_Li_water} %spread out from the present theoretical curves
for $E<600$~keV/amu. 
In the case of uracil, recent CTMC calculations by Sarkadi~\cite{sarkadi2016} are also above the experimental values by Tribedi group~\cite{agnihotri2012,agnihotri2013}.

@@

In our previous work, we noticed that the CDW ionization cross 
sections, $\sigma_A$, of atomic targets H, C, N, and O  scale with the number of
active electrons per atom $\nu_A$, as \mbox{$\sigma_e=\sigma_A/\nu_A$,} 
where $\nu_A$ is $1$ for H and $4$ for C, N, O, i.e.,
\begin{equation}
 \frac{\sigma_{\mathrm{H}}}{1}\sim
 \frac{\sigma_{\mathrm{C}}}{4}\sim
 \frac{\sigma_{\mathrm{N}}}{4}\sim
 \frac{\sigma_{\mathrm{O}}}{4}\,.
\end{equation}
By %considering
means of the SSM, we define the number of active electrons per
molecule as $n_e=\sum_A n_A \nu_A$. The $n_e$ values for the 
molecular targets considered throughtout this work are displayed in 
Table~\ref{nn}. The scaling with the molecular number of active
electrons proved to give excellent results, as shown in Fig.~6 of 
Ref.~\cite{MendezJPB20}.

@@

By incorporating the $Z^\alpha$ reduction and the scaling with the %molecular 
number  of active electrons% scaling
, we introduce the scaled and reduced
ionization cross section of molecules $\tilde{\sigma}$, which is 
expressed as a function of $E/Z^{2-\alpha}$, and it is given by
\begin{equation}
 \tilde{\sigma}=\frac{\sigma_e}{Z^{\alpha}}=\frac{\sigma_M/n_e}{Z^{\alpha}}\,,
\label{eq:u-scaling}
\end{equation}
where $\sigma_M$ is the ionization cross section for the molecular target, 
$n_e$ is the number of active electrons per molecule displayed in 
Table~\ref{nn} and the parameter is $\alpha=1.2$.
Fig.~\ref{fig:zalpha} shows the %reduced-scaled 
theoretical and 
experimental values of $ \tilde{\sigma}$ (given by Eq.~(\ref{eq:u-scaling})) %data 
for all the systems displayed in Fig.~\ref{fig:zreduced}% obtained using Eq.~(\ref{eq:u-scaling})
. As can be noted, the scaling works very well
and is independent of the ion charge or the complexity of the molecular
target. 
Our theoretical curves lay in a narrow band valid for any charged ion 
(reduced with $Z^\alpha$) in any molecule (scaled with the number of 
active electrons) with a dispersion of about 20\%. If we consider
the experimental data, the uncertainty of our scaling grows to 30\%, 
which is coverd by %shown with 
the gray area in Fig.~\ref{fig:zalpha}.
It is worth noting that we did not include in this figure %Fig.~\ref{fig:zalpha}
the data for uracil from Refs.~\cite{agnihotri2012,agnihotri2013}, 
and for Li$^{+3}$ on water~\cite{Luna_Li_water}. The discussion 
about these experimental values exceeds the present work.

\begin{figure*}[!htb]%[h]
\centering
\includegraphics[width=0.75\textwidth]{zmol_werror.eps}
\caption{(Color online) Ionization cross section reduced with the ion
charge $Z$ and scaled with number of active electrons per molecule $n_e$,
given by Eq.~(\ref{eq:u-scaling}) with $\alpha=1.2$. 
Curves: present CDW-SSM theoretical results. 
Symbols: experimental impact of H$^+$ on 
\mbox{\LARGE$\color{red}\circ$} adenine~\cite{iriki2011}, 
{\fontsize{11}{20}$\color{red}\pmb{\triangle}$} uracil~\cite{itoh2013}, 
{\fontsize{11}{20}$\color{red}\pmb{\triangledown}$} pyrimidine~\cite{wolff2014} and 
{\fontsize{10}{20}$\color{red}\pmb{\meddiamond}$} THF~\cite{wang2016};
{\fontsize{11}{20}$\color{Purple}\pmb{\lozenge}$} C$^{+6}$ on adenine \cite{tribedi2019};
\mbox{\fontsize{11}{20}$\color{red}\pmb{\odot}$}~\cite{Luna2007}, 
{\fontsize{11}{20}\color{red}$\pmb{\logof}$}~\cite{Rudd86}, 
{\fontsize{11}{20}\color{red}$\pmb{\varowedge}$}~\cite{pRudd85}, 
{\fontsize{11}{20}\color{red}$\pmb{\varoslash}$}~\cite{toburen80} H$^+$,
{\fontsize{11}{20}\color{ForestGreen}$\pmb{\varolessthan}$}~\cite{Ohsawa05},
{\fontsize{11}{20}\color{ForestGreen}$\pmb{\varogreaterthan}$}~\cite{Rudd85},
{\fontsize{11}{20}\color{ForestGreen}$\pmb{\varoslash}$}~\cite{toburen80} He$^{+2}$,
{\fontsize{11}{20}\color{Purple}$\pmb{\ovee}$}~C$^{+6}$ \cite{DalCappello2009,Bhattacharjee17}, and 
{\fontsize{11}{20}\color{BurntOrange}$\pmb{\obslash}$}
O$^{+8}$ \cite{Tribedi_O_water} on water.
H$^{+}$ impact on 
{\fontsize{11}{20}${\color{red}\pmb{\boxast}}$}~N$_2$, 
{\fontsize{11}{20}${\color{red}\pmb{\boxbox}}$}~O$_2$, 
{\fontsize{11}{20}${\color{red}\pmb{\boxbar}}$}~CO, 
{\fontsize{11}{20}${\color{red}\pmb{\boxbslash}}$}~CO$_2$, and
{\fontsize{11}{20}${\color{red}\pmb{\boxdot}}$}~CH$_4$; 
and He$^{+2}$ impact on 
{\fontsize{11}{20}${\color{ForestGreen}\pmb{\boxast}}$}~N$_2$,
{\fontsize{11}{20}${\color{ForestGreen}\pmb{\boxbox}}$}~O$_2$, 
{\fontsize{11}{20}${\color{ForestGreen}\pmb{\boxbar}}$}~CO, 
{\fontsize{11}{20}${\color{ForestGreen}\pmb{\boxbslash}}$}~CO$_2$, and
{\fontsize{11}{20}${\color{ForestGreen}\pmb{\boxdot}}$} 
CH$_4$~\cite{Rudd85,Rudd1983}, 
{\fontsize{11}{20}${\color{red}\pmb{\varowedge}}$}~H$^{+}$ on 
CH$_4$~\cite{Luna2019}; 
and electron impact on $\rhd$~pyrimidine~\cite{bug2017}, and $\lhd$, 
$\medstar$~\cite{wolf2019,fuss2009} THF.}
\label{fig:zalpha}
\end{figure*}

We consider this scaling robust enough to be valid for different 
ion--molecule combinations. We tested the generality of our model 
by including in Fig.~\ref{fig:zalpha} several data sets of 
molecular targets not considered previously, such as the measurements 
by Rudd~\textit{et al.}~\cite{Rudd85,Rudd1983} for H$^{+}$ and He$^{+2}$ 
in N$_2$, O$_2$, CH$_4$, CO and CO$_2$, and the recent values by 
Luna~\textit{et al.} \cite{Luna2019} for H$^{+}$ in CH$_4$.

The good agreement shown in Fig.~\ref{fig:zalpha} summaries the main 
result of this work, and holds the validity of the present scaling for different ions and targets.
%with the ion charge and the number of active electrons in the target. 
Although the theoretical CDW-SSM results are valid for energies above 
the maximum of the cross sections, %it is worth noting from Fig.~\ref{fig:zalpha} that 
the scaling of the experimental data extends 
even to lower impact energies, as can be noted in Fig.~\ref{fig:zalpha}. 
New measurements for other ions and molecules are expected to 
reinforce the present proposal.

@@

% our ---------------------------------------------------------
\bibitem{MendezJPB20}
A. M. P. Mendez, C. C. Montanari, and J. E. Miraglia, 
J. Phys. B: At. Mol. Opt. Phys.  \textbf{53}, 055201 (2020).

@@

\bibitem{itoh2013} 
A. Itoh, Y. Iriki, M. Imai, C. Champion, and R. D. Rivarola, 
Phys. Rev. A \textbf{88}, 052711 (2013).

\bibitem{iriki2011}
Y. Iriki, Y. Kikuchi, M. Imai, and A. Itoh
Phys. Rev. A \textbf{84}, 052719 (2011).

>>>>

\definecolor{olive}{RGB}{182,187,37}
\usepackage[mathlines]{lineno}

@@

\title[Scaling rules for the ionization of biological
molecules]{Scaling rules for the ionization of biological
molecules by highly charged ions}
\author{A. M. P. Mendez, C. C. Montanari, J. E. Miraglia}
\affiliation{Instituto de Astronom\'{\i}a y F\'{\i}sica del Espacio, 
Consejo Nacional de Investigaciones Cient\'ificas y T\'ecnicas - 
Universidad de Buenos Aires, Pabll\'on IAFE, 1428 Buenos Aires, 
Argentina}

\begin{abstract}
We investigate scaling rules for the ionization cross sections of
multicharged ions on molecules of biological interest. The cross 
sections are obtained using a methodology presented in [Mendez 
\textit{et al.} J. Phys B (2020)], which considers distorted-wave 
calculations for atomic targets combined with a molecular 
stoichiometric model. We examine ions with nuclear charges $Z$ from 
$+1$ to $+8$ impacting on five nucleobases --adenine, cytosine, 
guanine, thymine, uracil--, tetrahydrofuran, pyrimidine, and water. 
We investigate scaling rules of the ionization cross section with 
the ion charge and the number of active electrons per molecule. 
Combining these two features, we define a scaling law for any ion and 
molecular target, which is valid in the intermediate to high energy 
range, i.e., 0.2-5 MeV/amu for oxygen impact. Thus, the forty 
ion-molecule systems analyzed here can be merged into a single band. 
We confirm the generality of our independent scaling law with several
collisional systems. 
\end{abstract}

@@

\maketitle

@@

This contribution constitutes a follow-up to our previous 
work~\cite{MendezJPB20}. We introduce here a two-folded scaling rule 
for the ionization cross sections of complex molecules by charged ions. 
Our approach considers the dependence of the cross section with the ion 
charge $Z$ and incorporates the scaling of the ionization with the 
number of active electron $n_e$ of the molecular targets. Scaling rules 
are generally very useful since they can be used as first-order 
approximations in experimental measurements and multipurpose codes.

@@

\begin{figure*}[!htb]
\centering
\includegraphics[width=0.9\textwidth]{zscale_alpha-wref.eps}
\caption{(Color online) Scaled ionization cross section $\sigma/Z^{\alpha}$ 
as a function of the ion impact energy $E/Z^{2-\alpha}$ with $\alpha=1.2$. 
Colors are associated with the incident ion labeled on top of the figure. 
Curves: present CDW-SSM theoretical results. 
Symbols: experimental data~\cite{itoh2013,iriki2011,wolff2014,wang2016,
tribedi2019,agnihotri2012,agnihotri2013,Luna2007,Rudd86,pRudd85,
toburen80,Ohsawa05,Bhattacharjee17,Luna_Li_water,DalCappello2009,
Tribedi_O_water}. Electron impact ionization values~\cite{rahman2016,
bug2017,wolf2019,fuss2009} are included with the corresponding 
equi-velocity conversion.}
\label{fig:zreduced}
\end{figure*}

We found two types of $Z$-scaling laws in the literature applicable 
to the intermediate impact energy range. The rule suggested by Janev 
and Presnyakov~\cite{janev1980} considers $\sigma/Z$ versus $E/Z$ to be 
the \textit{natural} reduced form of the ionization cross section 
$\sigma$ and the incident ion energy $E$. More recently, Montenegro and 
co-workers~\cite{dubois13,montenegro_pra13} suggested an alternative 
scaling by taking into account that the cross section is a function of 
$Z^2/E$ at high energies. Their scaling, given by 
\begin{equation}
 \sigma/Z^{\alpha}=f(E/Z^{2-\alpha}),
\label{eq:Montenegro}
\end{equation}
keeps the $Z^2/E$ relationship for any value of the parameter 
$\alpha$. The authors proposed $\alpha=4/3$ for the ionization of He and 
H$_2$ by differently charged ions~\cite{dubois13}.

@@

We also examined the experimental data available for the forty 
ion-target systems~\cite{itoh2013,iriki2011,wolff2014,wang2016,
tribedi2019,agnihotri2012,agnihotri2013,Luna2007,Rudd86,pRudd85,
toburen80,Ohsawa05,Bhattacharjee17,Luna_Li_water,DalCappello2009,
Tribedi_O_water} with the $Z^\alpha$-scaling rule. For targets with 
none or little experimental data, we included electron impact 
ionization results~\cite{rahman2016,bug2017,wolf2019,fuss2009} at high 
velocity with the corresponding equivelocity conversion. As can be 
noted, most of the data in Fig.~\ref{fig:zreduced} confirm the present 
scaling, even for O$^{+8}$ in water~\cite{Tribedi_O_water}. Only two 
data sets are off our predictions: the ionization cross section of 
uracil by swift C, O, and F ions from Refs.~\cite{agnihotri2012,
agnihotri2013}, and the values for Li$^{+3}$ in water from 
Ref.~\cite{Luna_Li_water} for $E<600$~keV/amu. In the case of uracil, 
recent CTMC calculations by Sarkadi~\cite{sarkadi2016} are also above 
the experimental values by Tribedi and collaborators~\cite{agnihotri2012,
agnihotri2013}.

@@

In our previous work, we noticed that the CDW ionization cross sections 
$\sigma_A$ of atomic targets H, C, N, and O scale with the number of 
active electrons per atom $\nu_A$, as \mbox{$\sigma_e=\sigma_A/\nu_A$,} 
where $\nu_A$ is $1$ for H and $4$ for C, N, O, i.e.,
\begin{equation}
 \frac{\sigma_{\mathrm{H}}}{1}\sim
 \frac{\sigma_{\mathrm{C}}}{4}\sim
 \frac{\sigma_{\mathrm{N}}}{4}\sim
 \frac{\sigma_{\mathrm{O}}}{4}\,.
\end{equation}
By means of the SSM, we define the number of active electrons per 
molecule as $n_e=\sum_A n_A \nu_A$. The $n_e$ values for the 
molecular targets considered throughout this work are displayed in 
Table~\ref{nn}. The scaling with the molecular number of active
electrons proved to give excellent results, as shown in Fig.~6 of 
Ref.~\cite{MendezJPB20}.

@@

By incorporating the $Z^\alpha$ reduction and the scaling with the 
number of active electrons, we introduce the scaled and reduced 
ionization cross section of molecules $\tilde{\sigma}$, which is 
expressed as a function of $E/Z^{2-\alpha}$, and it is given by
\begin{equation}
 \tilde{\sigma}=\frac{\sigma_e}{Z^{\alpha}}=\frac{\sigma_M/n_e}{Z^{\alpha}}\,,
\label{eq:u-scaling}
\end{equation}
where $\sigma_M$ is the ionization cross section for the molecular 
target, $n_e$ is the number of active electrons per molecule displayed 
in Table~\ref{nn}, and the parameter is $\alpha=1.2$. 
Fig.~\ref{fig:zalpha} shows the theoretical and experimental values of 
$\tilde{\sigma}$ (given by Eq.~\ref{eq:u-scaling}) for all the systems 
displayed in Fig.~\ref{fig:zreduced}. As can be noted, the scaling 
works very well and is independent of the ion charge or the complexity 
of the molecular target. Our theoretical curves lay in a narrow band 
valid for any charged ion (reduced with $Z^\alpha$) in any molecule 
(scaled with the number of active electrons) with a dispersion of about 
$\pm 20\%$. If we consider the experimental data, the uncertainty of 
our scaling grows to $\pm 30\%$, which is schematized in 
Fig.~\ref{fig:zreduced} with a gray area. It is worth noting that we 
did not include in this figure the data for uracil from 
Refs.~\cite{agnihotri2012,agnihotri2013}, and Li$^{+3}$ on 
water~\cite{Luna_Li_water}. The discussion about these experimental 
values exceeds the present work.

\begin{figure*}[!htb]
\centering
\includegraphics[width=0.8\textwidth]{zmol_werror-wref.eps}
\caption{(Color online) Ionization cross section reduced with the ion
charge $Z$ and scaled with number of active electrons per molecule $n_e$,
given by Eq.~(\ref{eq:u-scaling}) with $\alpha=1.2$. 
Curves: present CDW-SSM theoretical results. 
Symbols: experimental data~\cite{itoh2013,iriki2011,wolff2014,wang2016,
tribedi2019,Luna2007,Rudd86,pRudd85,toburen80,Ohsawa05,Bhattacharjee17,
DalCappello2009,Tribedi_O_water,Rudd85,Luna2019,Rudd1983}. Electron 
impact ionization values~\cite{rahman2016,bug2017,wolf2019,fuss2009} 
are included with the corresponding equi-velocity conversion.}
\label{fig:zalpha}
\end{figure*}

We consider the present scaling robust enough to be valid for different 
ion-molecule combinations. We tested the generality of our model 
by including in Fig.~\ref{fig:zreduced} several data sets of 
molecular targets not considered previously, such as the measurements 
by Rudd~\textit{et al.}~\cite{Rudd85,Rudd1983} for H$^{+}$ and He$^{+2}$ 
in N$_2$, O$_2$, CH$_4$, CO and CO$_2$, and recent values by 
Luna~\textit{et al.} \cite{Luna2019} for H$^{+}$ in CH$_4$ .

The good agreement shown in Fig.~\ref{fig:zalpha} summarizes the main 
result of this work and holds the validity of the present scaling for 
different ions and targets. Although the theoretical CDW-SSM results 
are valid for energies above the maximum of the cross sections, the 
scaling of the experimental data extends even to lower impact energies, 
as can be noted in Fig.~\ref{fig:zalpha}. New measurements for other 
ions and molecules are expected to reinforce the present proposal.

@@

% ours ---------------------------------------------------------
\bibitem{MendezJPB20}
A. M. P. Mendez, C. C. Montanari, and J. E. Miraglia, 
J. Phys. B: At. Mol. Opt. Phys.  \textbf{53}, 055201 (2020).

@@

%17
\bibitem{iriki2011}
Y. Iriki, Y. Kikuchi, M. Imai, and A. Itoh
Phys. Rev. A \textbf{84}, 052719 (2011). 
%18
\bibitem{tribedi2019} 
S. Bhattacharjee, C. Bagdia, M. R. Chowdhury, A. Mandal, J. M. Monti, 
R. D. Rivarola, and L. C. Tribedi, 
